Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2021-09-15 RR_manuscript/Draft_RR.tex
♻️ - fixes to RR

\begin{abstract}
Recent efforts to replicate published findings have uncovered surprisingly low success rates across disciplines. Moreover, several studies have highlighted the large degree of analytic flexibility in data analysis which can lead to substantially different conclusions based on the same data set. Thus, researchers have expressed their concerns that these researcher degrees of freedom might facilitate bias and can lead to claims that do not stand the test of time. Even greater flexibility is to be expected in fields in which the primary data lend themselves to a variety of possible operationalizations. The multidimensional, temporally extended nature of speech constitutes an ideal testing ground for assessing the variability in analytic approaches that derives not only from aspects of statistical modelling but also from decisions regarding the quantification of the measured behavior. In the present study, we will give the same speech production dataset to N teams of researchers and will ask them to answer the same research question. Using Bayesian meta analytical tools, we will quantify this variability and relate it to predictors related to the analysts and their analyses.
\end{abstract}

@@

In order to effectively accumulate knowledge, science needs (i) to produce data that can be replicated using the original methods and (ii) to arrive at robust conclusions substantiated by such data.
In recent coordinated efforts to replicate published findings, scientific disciplines have uncovered surprisingly low success rates \citep[e.g.,][]{open2015estimating, camerer2018evaluating} leading to what is now referred to as the \emph{replication crisis}.
Beyond the difficulties of replicating scientific findings, a growing body of evidence suggests that the conceptual conclusions are often variable even when researchers have access to the same data.
The latter situation has been referred to as the \emph{inference crisis} \citep{rotello2015more, starns2019assessing} and is, among other things, rooted in the inherent flexibility of data analysis \citep[often referred to as researcher degrees of freedom:][]{simmons2011false, gelman2014statistical}.
Data analysis involves many different steps, such as inspecting, organizing, transforming, and modeling the data, to name a few.
Along the way, different methodological and analytic choices need to be made, all of which may influence the final interpretation of the data.

These researcher degrees of freedom are both a blessing and a curse.
They are a blessing because they afford us the opportunity to look at nature from different angles, which, in turn, allows us to make important discoveries and generate new hypotheses \citep[e.g.,][]{box1976science, tukey1977exploratory, de2014thought}.
They are a curse because idiosyncratic choices can lead to categorically different interpretations, which eventually find their way into the publication record where they are taken for granted \citep{simmons2011false}.
Recent projects have shown that the variability between different data analysts is vast and can lead independent researchers to draw different conclusions from the same data set \citep[e.g.,][]{silberzahn2018many, starns2019assessing, botvinik2020variability}.
These studies, however, might still underestimate the extent to which analysts vary because data analysis is not restricted to the statistical analysis of ready-made numeric data.
These data can in fact be the result of complex measurement processes that translate a phenomenon, such as human behavior, into numbers.
This is particularly true for fields that draw conclusions about human behavior and cognition from multidimensional data like audio or video data.
In fields working on human speech production, for example, researchers need to make numerous decisions about what to measure and how to measure it, in other words, how to operationalise the phenomenon under study.
This is not trivial, given the temporal extension of the acoustic signal and its complex structural composition.

In this article, we investigate the impact of analytic choices on research results when many analyst teams analyze the same speech production data set, a process that involves both decisions regarding the \emph{operationalization} of a complex signal and decisions regarding \emph{statistical analysis}.
Specifically, we discuss the degree of variability in research results obtained by N teams who had to choose the operationalisation and statistical procedures to answer the same research question, on the basis of the same same set of speech recordings.
Our goals are twofold: First, our study conceptually replicates previous many-analyses projects, by probing the effects of different statistical analyses and by assessing the generalizability of published findings to other disciplines (here, the speech sciences).
Second, our study extends the scope of inquiry to include flexibility in the operationalisation of complex human behavior (here, human speech).
This is an important addition in that the increased number of ``forking paths'' in the ``garden of analytic choices'', deriving from the many decisions involved in quantification, might reveal a higher degree of variability across analysts than what previously observed.

@@

This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways of being measured \citep{roettger2019researcher}.
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of statistical training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions \citep{roettger2019emergent}.
Analytic flexibility has been widely discussed from a conceptual point of view \citep{simmons2011false, wagenmakers2012agenda, nosek2014method} and in regard to its application in individual scientific fields \citep[e.g.][]{wicherts2016, charles2019, roettger2019researcher}.
This notwithstanding, there are still many unknowns regarding the extent of analytic plurality in practice.

@@

Recent collaborative attempts have started to shed light on how different analysts tackle the same data set and have revealed a large amount of variability.
In a collaborative effort, \citet{silberzahn2018many} let twenty-nine independent analysis teams address the same research hypothesis.
Analytic approaches and consequently the results varied widely between teams.
Sixty-nine percent of the teams found support for the hypothesis, and 31\% did not.
Out of the 29 analytic strategies, there were 21 unique combinations of covariates.
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses.
The authors results suggest that analytic plurality may be an inevitable byproduct of the scientific process and not necessarily driven by different levels of expertise nor bias.

Several other recent studies corroborated this analytic flexibility across different disciplines.
\citet{dutilh2019} and \citet{starns2019assessing} investigated analysts' choices when inferring theoretical constructs based on the same data set using computational models.
Both studies revealed vastly different modeling strategies, even though scientific conclusions were similar across analysis teams (see also \citet{Parker2020} on analytic flexibility in ecology, and \citet{botvinik-nezer2020} on neuroimaging data).
\citet{bastiaansen2020} crowdsourced clinical recommendations based on analyses of an individual patient.
Results suggests that analysts differed substantially regarding decisions related to both the statistical analysis of the data and the theoretical rationale behind interpreting the statistical results.

Extending on the many-analysts approach, \citet{landy2020crowdsourcing} asked 15 research teams to independently design studies to answer the same research questions.
Again, they found vast heterogeneity in researchers' conclusions which was not predicted by the researchers' expertise but seem to be more dependent on the research question itself.
This is in line with a recent re-analysis of the study by \citet{silberzahn2018many} by \citet{auspurg2021has}.
The authors argue that the observed heterogeneity across analysts in \citet{silberzahn2018many} might be driven by flexibility in interpreting the research question.

While these studies attested a large degree of analytic flexibility with possibly impactful consequences, they only investigated analytic decisions related to statistical inference or the architecture of computational models.
In these studies the data sets were fixed and neither data collection nor measurement could be changed.
However, in many fields the primary raw data are complex signals that need to be operationalized relative to a theoretically motivated research question.
This is especially true in the social sciences, where the phenomenon under investigation corresponds to both observable and unobservable human behavior.

@@

In those disciplines concerned with communication, human behavior often corresponds to multidimensional visual or acoustic signals.
The complex nature of this data multiplies the number of possible analytic approaches, thus further increasing analytic flexibility.
The present study looks at experimentally elicited speech production data in order to understand how this flexibility manifests itself in a scenario where several procedures are involved not only in the statistical analysis of numeric data, but also in the operationalization and measuring of complex signals necessary to derive such data.

@@

Speech is a complex signal that is characterized by structurally different acoustic landmarks distributed throughout different temporal domains.
Thus, choosing how to measure a phenomenon of interest is an important analytic decision.
Take for example the following sentence in (1).

@@

The multiplicity of phonetic measurements grows exponentially if we look at larger temporal domains as is the case for suprasegmental aspects of speech.
For example, studies investigating acoustic correlates of word stress (e.g.~the difference between \emph{ínsight} and \emph{incíte}) use a wide variety of measurements, including temporal characteristics (duration of certain segments or sub-segmental intervals), spectral characteristics (intensity, formants, and spectral tilt), and measurements related to fundamental frequency (f0) \citep[e.g.,][]{gordon2017acoustic}.
Moving on to the expression of higher-level functions, like information structure and discourse pragmatics, relevant acoustic cues can be distributed throughout even larger domains, such as phrases and whole utterances \citep[e.g.,][]{ladd2008intonational}.
Differences in position, shape, and alignment of pitch modulations over multiple locations within a sentence are correlated with differences in discourse functions \citep[e.g.,][]{niebuhr2011}.
The latter can also be expressed by global vs local pitch modulations \citep{heuven2002}, as well as acoustic information within the temporal or spectral domain \citep[e.g.,][]{van2005speech}.
Extra-linguistic information, like the speaker's intentions, levels of emotional arousal or social identity, are also conveyed by broad-domain parameters, such as voice quality, rhythm, and pitch \citep{foulkes2006, ogden2004, white2009}.

In short, when testing hypotheses on speech production data, researchers are faced with many choices and possibilities.
The larger the functional domain (e.g.~segments vs words vs utterances), the higher the number of conceivable operationalizations.
For example, several decisions have to be made when comparing the two realizations of the sentence in (1), one of which is intended to signal emphasis on \emph{another} and one of which emphasizes \emph{Zoom}.

@@

Do we compare only the word \emph{another} in (2a) and (2b), or also the word \emph{Zoom}?
Do we measure utterance-wide acoustic profiles, whole words, or just the stressed syllables?
Do we average across the chosen time domain or do we measure a specific point in time?
Do we measure fundamental frequency, intensity, or something else \citep{stevens2000}?

@@

When looking at phrase-level temporal domains, the number of alternative analytic pipelines increases substantially.
Figure \ref{fig:forkingPaths}A shows a typical example of a decision tree with which speech researchers are often confronted.
Each of the four analytic decisions in the example have different possible options.
Here only one particular path has been taken, and a different one would likely produce quite different results.
Once we have decided to compare fundamental frequency of the word \emph{another} across the two utterances, there are still many choices to be made, all of which need to be justified.
As Figures \ref{fig:forkingPaths}B-C illustrate, we could measure f0 at specific points in time like the onset of the temporal window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data in previous work \citep[see][]{gordon2017acoustic}.
But the journey in the garden of analytic paths goes on.
Other important operationalization steps could involve filtering the audio signal, smoothing the extracted f0 track, removing values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on.

These decisions are intended to be made prior to any statistical analysis, but are at times revised a posteriori in light of unforeseen or surprising outcomes (i.e.~after data collection and/or preliminary analyses).
These myriads of possible decisions are multiplied by those researcher degrees of freedom related to statistical analysis \citep[e.g.][]{wicherts2016}.
In sum, speech data is made of complex physical signals that generates an as-of-yet unappreciated amount of analytic flexibility in the choice of measures and operationalizations.
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines, including both operationalization and statistical analysis, across independent researchers, we provided analytic teams with an experimentally elicited speech production data set.
The data set derives from the unpublished research project \emph{Prosodic encoding of redundant referring expressions}, which set out to investigate whether speakers acoustically modify utterances to signal unexpected referring expressions.\footnote{Results of this research project were neither published nor publicly presented and are stored on a private OSF repository.}
In the following section we introduce the research question and the experimental procedure of said project, and we describe the resulting data set as used in the current study.

@@

Referring is one of the most basic and prevalent uses of language and one of the most widely researched areas in language science.
It is an open question how speakers choose a referring expression when they want to refer to a specific entity like a banana.
The context within which an entity occurs (i.e., with other non-fruits, other fruits, or other bananas) plays a large part in determining the choice of referring expressions.
Generally, speakers aim to be as informative as possible to uniquely establish reference to the intended object, but they are also resource-efficient in that they avoid redundancy \citep{grice1975logic}.
Thus one would expect the use of a modifier, for example, only if it is necessary for disambiguation.
For instance, one might use the adjective \emph{yellow} to describe a banana in a situation in which there is a yellow and a less ripe green banana available, but not when there is only one banana to begin with.

Despite the coherent idea that speakers are both rational and efficient, there is much evidence that speakers are often over-informative.
Speakers use referring expressions that are more specific than strictly necessary for the unambiguous identification of the intended referent \citep{sedivy2003pragmatic, westerbeek2015stored, rubio2016redundant}, which has been argued to facilitate object identification and make communication between speakers and listeners more efficient \citep{arts2011overspecification, paraboni2007generating, rubio2016redundant}.
Recent findings suggest that the utility of a referring expression depends on how good it is for a listener (compared to other referring expressions) to identify a target object.
For example, \citet{degen2020redundancy} showed that modifiers that are less typical for a given referent (e.g.~a blue banana) are more likely to be used in an over-informative scenario (e.g.~when there is just one banana).
This account, however, has mainly focused on content selection \citep{gatt2013we}.

Even when morphosyntactically identical expressions are involved, speakers can modulate utterances via suprasegmental acoustic properties like temporal and spectral modifications \citep[e.g.,][]{ladd2008intonational}.
Most prominently, languages use intonation to signal discourse relationships between referents, among other functions.
Intonation marks discourse-relevant referents for being new or given information, to guide the listeners' interpretation of incoming messages.
Beyond structuring information relative to the discourse, a few studies suggested that speakers might use intonation to signal atypical lexical combinations \citep[e.g.][]{dimitrova2008prosodic, dimitrova2009did}.
Referential expressions such as \emph{blue banana} were produced with greater prosodic prominence than more typical referents such as \emph{yellow banana}.
These results are in line with the idea of resource-efficient, rational language users who modulate their speech in order to facilitate listeners' comprehension.
However, the above studies are based on a rather small sample size (10 participants) and on potentially anti-conservative statistical analyses, leaving reason to doubt the generalizability of the studies' conclusions.

To further illuminate the question of whether speakers modify speech to signal atypical referents, and overcome some of the limitations of previous work, thirty native German speakers were recorded in a production study while interacting with a confederate (one of the experimenters) in a referential game, following expermintal procedures typical of the field.
The participants had to verbally instruct the confederate to select a specified target object out of four objects presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two computer screens.
The participant and the experimenter could not see each other nor each others' screens.
Figure \ref{fig:procedural} shows the experimental procedure time-line.
After a familiarization phase, the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g.~\emph{Gelbe Zitrone} `yellow lemon', \emph{Rote Gurke} `red cucumber', \emph{Rote Socken} `red socks').

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
Focused units denote the set of all (contextually relevant) alternatives \citep[e.g.][]{rooth1992theory}.
Concretely, a focus contrast marks one of more elements in a sentence as prominent, by different linguistic means depending on the language.
For instance, if the competitor and target objects differ but their color did not (e.g.~\emph{yellow banana} vs \emph{yellow tomato}), the noun is said to be in focus (the Noun Focus condition, NF).
If the objects are the same but differ in color (e.g.~\emph{yellow banana} vs \emph{blue banana}), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g.~\emph{yellow banana} vs \emph{blue tomato}), the the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g.~\emph{orange mandarin}), medium typical (e.g.~\emph{green tomato}), or atypical (e.g.~\emph{yellow cherry}), as established by a norming study that was conducted prior to the production experiment just described.\footnote{A detailed description of the norming and production studies from the \emph{Prosodic encoding of redundant referring expressions} project, which was given to the analysts with the data set, can be found in \texttt{methods\_norm\_prod.pdf} at (temporary link) \url{https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f}}
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 (\(15 \times 2 \times 30\) participants) spoken utterances.

For the present study, N analysis teams will receive access to the entire data set generated by the production study.
The data set is constituded by audio recordings and annotation files in a format that is typical for the field.
The teams will be instructed to answer the following research question, using the provided data set: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}

@@

We estimate that this process, from the time of an in-principle acceptance of this Stage 1 Registered Report to the end of Phase 5, will take nine months.
The factor most likely to delay our time-line is the rate of completion of the original set of analyses by independent analysis teams.
The following sections report the criteria for sample size, data exclusions, data manipulations, and all the measures in the study.

@@

The analysis teams will register for participation and each of the analysts individually will answer a demographic and expertise questionnaire (\texttt{intake\_form.pdf}).
PDF versions of all the questionnaires used in this study are available at LINK\footnote{Temporary link: \url{https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f}, Questionnaires component.}.
The questionnaire collects information on the analysts current position and self-estimated breadth and level of statistical expertise and acoustic analysis skills.
We will then request that they answer the question: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
To do so, they will be given the data generated by the experiment described in Section \protect\hyperlink{s:dataset}{The data set}.
The data will include the audio recordings with corresponding time-aligned transcriptions in the form of Praat TextGrid files (LINK\footnote{Temporary link: \url{https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f}, Data component.}).

>>>>

\begin{abstract}
Recent efforts to replicate published findings have uncovered surprisingly low success rates across disciplines. Moreover, several studies have highlighted the large degree of analytic flexibility in data analysis which can lead to substantially different conclusions based on the same data set. Thus, researchers have expressed their concerns that these researcher degrees of freedom might facilitate bias and can lead to claims that do not stand the test of time. Even greater flexibility is to be expected in fields in which the primary data lend themselves to a variety of possible operationalizations. The multidimensional, temporally extended nature of speech constitutes an ideal testing ground for assessing the variability in analytic approaches that derives not only from aspects of statistical modelling but also from decisions regarding the quantification of the measured behavior. In the present study, we will give the same speech production dataset to N teams of researchers and will ask them to answer the same research question. Using Bayesian meta-analytic tools, we will quantify this variability and relate it to predictors related to the analysts and their analyses.
\end{abstract}

@@

In order to effectively accumulate knowledge, science needs (i) to produce data that can be replicated using the original methods and (ii) to arrive at robust conclusions substantiated by such data.
In recent coordinated efforts to replicate published findings, scientific disciplines have uncovered surprisingly low success rates \citep[e.g.,][]{open2015estimating, camerer2018evaluating} leading to what is now referred to as the \emph{replication crisis}.
Beyond the difficulties of replicating scientific findings, a growing body of evidence suggests that researchers' conclusions often vary even when they have access to the same data.
The latter situation has been referred to as the \emph{inference crisis} \citep{rotello2015more, starns2019assessing} and is, among other things, rooted in the inherent flexibility of data analysis \citep[often referred to as researcher degrees of freedom:][]{simmons2011false, gelman2014statistical}.
Data analysis involves many different steps, such as inspecting, organizing, transforming, and modeling the data, to name a few.
Along the way, different methodological and analytic choices need to be made, all of which may influence the final interpretation of the data.

These researcher degrees of freedom are both a blessing and a curse.
They are a blessing because they afford us the opportunity to look at nature from different angles, which, in turn, allows us to make important discoveries and generate new hypotheses \citep[e.g.,][]{box1976science, tukey1977exploratory, de2014thought}.
They are a curse because idiosyncratic choices can lead to categorically different interpretations, which eventually find their way into the publication record where they are taken for granted \citep{simmons2011false}.
Recent projects have shown that the variability between different data analysts is vast and can lead independent researchers to draw different conclusions from the same data set \citep[e.g.,][]{silberzahn2018many, starns2019assessing, botvinik2020variability}.
These studies, however, might still underestimate the extent to which analysts vary because data analysis is not restricted to the statistical analysis of ready-made numeric data.
These data can in fact be the result of complex measurement processes that translate a phenomenon, such as human behavior, into numbers.
This is particularly true for fields that draw conclusions about human behavior and cognition from multidimensional data like audio or video data.
In fields working on human speech production, for example, researchers need to make numerous decisions about what to measure and how to measure it, in other words, how to operationalise the phenomenon under investigation.
This is not trivial, given the temporal extension of the acoustic signal and its complex structural composition.

In this article, we investigate the impact of analytic choices on research results when many analyst teams analyze the same speech production data set, a process that involves both decisions regarding the \emph{operationalization} of a complex signal and decisions regarding \emph{statistical analysis}.
Specifically, we discuss the degree of variability in research results obtained by N teams who had to choose the operationalisation and statistical procedures to answer the same research question, on the basis of the same set of raw data (here, speech recordings).
Our goals are twofold: (i) our study conceptually replicates previous many-analyses projects, by probing the effects of different statistical analyses and by assessing the generalizability of published findings to other disciplines (here, the speech sciences);
(ii) our study extends the scope of inquiry to include flexibility in the operationalisation of complex human behavior (here, human speech).
This is an important addition in that the increased number of ``forking paths'' in the ``garden of analytic choices'', deriving from the many decisions involved in quantification, might reveal a higher degree of variability across analysts than what previously observed, thus giving us a more realistic estimate of variability across analysts.

@@

This issue is particularly amplified in fields in which the raw data lend themselves to many possible ways of being measured \citep{roettger2019researcher}.
Combined with a wide variety of methodological and theoretical traditions as well as varying levels of quantitative training across subfields, the inherent flexibility of data analysis might lead to a vast plurality of analytic approaches that can lead to different scientific conclusions \citep{roettger2019emergent}.
Analytic flexibility has been widely discussed from a conceptual point of view \citep{simmons2011false, wagenmakers2012agenda, nosek2014method} and in regard to its application in individual scientific fields \citep[e.g.][]{wicherts2016, charles2019, roettger2019researcher}.
This notwithstanding, there are still many unknowns regarding the extent of analytic plurality in practice.

@@

Recent collaborative attempts have started to shed light on how different analysts tackle the same data set and have revealed a large amount of variability.
In a seminal collaborative effort, \citet{silberzahn2018many} let twenty-nine independent analysis teams address the same research hypothesis.
Analytic approaches and consequently the results varied widely between teams.
Sixty-nine percent of the teams found support for the hypothesis, and 31\% did not.
Out of the 29 analytic strategies, there were 21 unique combinations of covariates.
Importantly, the observed variability was neither predicted by the team's preconceptions about the phenomenon under investigation nor by peer ratings of the quality of their analyses.

The authors results suggest that analytic plurality may be an inevitable byproduct of the scientific process and not necessarily driven by different levels of expertise or bias.

Several other recent studies corroborated this analytic flexibility across different disciplines.
\citet{dutilh2019} and \citet{starns2019assessing} investigated analysts' choices when inferring theoretical constructs based on the same data set using computational models.
Both studies revealed vastly different modeling strategies, even though scientific conclusions were similar across analysis teams (see also \citet{Parker2020} on analytic flexibility in ecology, and \citet{botvinik-nezer2020} on neuroimaging data).
\citet{bastiaansen2020} crowdsourced clinical recommendations based on analyses of an individual patient.
Their results suggests that analysts differed substantially regarding decisions related to both the statistical analysis of the data and the theoretical rationale behind interpreting the statistical results.

Extending on the many-analysts approach, \citet{landy2020crowdsourcing} asked 15 research teams to independently design studies to answer the same research questions.
Again, they found vast heterogeneity in researchers' conclusions which was not predicted by the researchers' expertise but seem to be more dependent on the research question itself.
This is in line with a recent re-analysis of \citet{silberzahn2018many} by \citet{auspurg2021has}.
The authors argue that the observed heterogeneity across analysts in \citet{silberzahn2018many} might have been driven by flexibility in statistically interpreting the research question.

While these studies attested a large degree of analytic flexibility with possibly impactful consequences, they focused on analytic decisions related to statistical inference or the architecture of computational models.
In these studies the data sets were fixed and neither data collection nor measurement could be changed.
Thus the estimates of variability found in the literature might reflect a lower bound only, ignoring large parts of the forking path related to measurement.

However, in many fields the primary raw data are complex signals that need to be operationalized relative to a theoretically motivated research question.
This is especially true in the social sciences, where the phenomenon under investigation corresponds to both observable and unobservable human behavior.

@@

In those disciplines concerned with communication, human behavior often corresponds to multidimensional visual and/or acoustic signals.
The complex nature of this data exponentiates the number of possible analytic approaches, thus further increasing analytic flexibility.
In order to estimate this increased flexibility, the present study looks at experimentally elicited speech production data.\\

@@

Speech is a complex signal that is characterized by structurally different acoustic landmarks distributed throughout different temporal domains.
Thus, choosing how to measure a communictive intention of interest is an important analytic step.
Take for example the following sentence in (1).

@@

The multiplicity of phonetic measurements grows exponentially if we look at larger temporal domains as is the case for suprasegmental aspects of speech.
For example, studies investigating acoustic correlates of word stress (e.g.~the difference between \emph{ínsight} and \emph{incíte}) use a wide variety of measurements, including temporal characteristics (duration of certain segments or sub-segmental intervals), spectral characteristics (intensity, formants, and spectral tilt), and measurements related to fundamental frequency (f0) \citep[e.g.,][]{gordon2017acoustic}.
Moving on to the expression of higher-level communicative functions, like information structure and discourse pragmatics, relevant acoustic cues can be distributed throughout even larger domains, such as phrases and whole utterances \citep[e.g.,][]{ladd2008intonational}.
Differences in position, shape, and alignment of pitch modulations over multiple locations within a sentence are correlated with differences in discourse functions \citep[e.g.,][]{niebuhr2011}.
The latter can also be expressed by global vs local pitch modulations \citep{heuven2002}, as well as acoustic information within the temporal or spectral domain \citep[e.g.,][]{van2005speech}.
Extra-linguistic information, like the speaker's intentions, levels of emotional arousal or social identity, are also conveyed by broad-domain parameters, such as voice quality, rhythm, and pitch \citep{foulkes2006, ogden2004, white2009}.

In short, when testing hypotheses on speakers' intentions with speech production data, researchers are faced with many choices and possibilities.
The larger the functional domain (e.g.~segments vs words vs utterances), the higher the number of conceivable operationalizations.
For example, several decisions have to be made when comparing the two realizations of the sentence in (1), one of which is intended to signal emphasis on \emph{another} and one of which emphasizes \emph{Zoom} (see 2a and 2b).

@@

Do we compare only the word \emph{another} in (2a) and (2b), or also the word \emph{Zoom}?
Do we measure utterance-wide acoustic profiles, whole words, or just stressed syllables?
Do we average across the chosen time domain or do we measure a specific point in time?
Do we measure fundamental frequency, intensity, or something else \citep{stevens2000}?

@@

When looking at phrase-level temporal domains, the number of alternative analytic pipelines increases substantially.
Figure \ref{fig:forkingPaths}A shows a typical example of a decision tree with which speech researchers are often confronted.
Each of the four analytic decisions in the example have different possible options.
Here only one particular path has been taken. A different one would likely produce different results and might lead to different conclusions.
Once we have decided to compare fundamental frequency of the word \emph{another} across the two utterances, there are still many choices to be made, all of which need to be justified.
As Figures \ref{fig:forkingPaths}B-C illustrate, we could measure f0 at specific points in time like the onset of the temporal window, the offset, or the midpoint.
We could also measure the value or time of the minimum or maximum f0 value.
We could summarize f0 across the entire window and extract the mean, median or standard deviation of f0, all of which have been used to analyze speech data in previous work \citep[see][]{gordon2017acoustic}.
But the journey in the garden of analytic paths goes on.
Other important operationalization steps could involve filtering the audio signal, smoothing the extracted f0 track, removing values that substantially deviate from surrounding values or expectations, either manually or automatically, and so on.

These decisions are intended to be made prior to any statistical analysis, but are at times revised a posteriori in light of unforeseen or surprising outcomes (i.e.~after data collection and/or preliminary analyses).
These myriads of possible decisions are multiplied by those researcher degrees of freedom related to statistical analysis \citep[e.g.][]{wicherts2016}.

In sum, speech data is made of complex physical signals that generates an as-of-yet unappreciated amount of analytic flexibility in the choice of measures and operationalizations.
The present paper probes this garden of forking paths in the analysis of speech.
To assess the variability in data analysis pipelines, including both operationalization and statistical analysis, across independent researchers, we provide analytic teams with an experimentally elicited speech production data set.
The data set derives from the unpublished research project \emph{Prosodic encoding of redundant referring expressions}, which set out to investigate whether speakers acoustically modify utterances to signal unexpected referring expressions.\footnote{Results of this research project were neither published nor publicly presented and are stored on a private OSF repository.}
In the following section we introduce the research question and the experimental procedure of said project, and we describe the resulting data set as used in the current study.

@@

Referring is one of the most basic and prevalent uses of language and one of the most widely researched areas in language science.
When trying to refer to a banana, what does a speaker say and how do they say it in a given context.
The context within which an entity occurs (i.e., with other non-fruits, other fruits, or other bananas) plays a large part in determining the choice of referring expressions.
Generally, speakers aim to be as informative as possible to uniquely establish reference to the intended object, but they are also resource-efficient in that they avoid redundancy \citep{grice1975logic}.
Thus one would expect the use of a modifier, for example, only if it is necessary for disambiguation.
For instance, one might use the adjective \emph{yellow} to describe a banana in a situation in which there are a yellow and a less ripe green banana available, but not when there is only one banana to begin with.

Despite the coherent idea that speakers are both rational and efficient, there is much evidence that speakers are often over-informative.
Speakers use referring expressions that are more specific than strictly necessary for the unambiguous identification of the intended referent \citep{sedivy2003pragmatic, westerbeek2015stored, rubio2016redundant}, which has been argued to facilitate object identification and make communication between speakers and listeners more efficient \citep{arts2011overspecification, paraboni2007generating, rubio2016redundant}.
Recent findings suggest that the utility of a referring expression depends on how good it is for a listener (compared to other referring expressions) to identify a target object.
For example, \citet{degen2020redundancy} showed that modifiers that are less typical for a given referent (e.g.~a blue banana) are more likely to be used in an over-informative scenario (e.g.~when there is just one banana).
This account, however, has mainly focused on content selection \citep{gatt2013we}, i.e.~what words to use.

Even when morphosyntactically identical expressions are involved, speakers can modulate utterances via suprasegmental acoustic properties like temporal and spectral modifications \citep[e.g.,][]{ladd2008intonational}.
Most prominently, languages can use intonation to signal discourse relationships between referents.
Intonation marks discourse-relevant referents for being new or given information, to guide the listeners' interpretation of incoming messages.
Beyond structuring information relative to the discourse, a few studies suggested that speakers might use intonation to signal atypical lexical combinations \citep[e.g.][]{dimitrova2008prosodic, dimitrova2009did}.
Referential expressions such as \emph{blue banana} were produced with greater prosodic prominence than more typical referents such as \emph{yellow banana}.
These results are in line with the idea of resource-efficient, rational language users who modulate their speech in order to facilitate listeners' comprehension.
However, the above studies are based on a small sample size (10 participants) and on potentially anti-conservative statistical analyses, leaving reason to doubt the generalizability of the studies' conclusions.

To further illuminate the question of whether speakers modify speech to signal atypical referents, and overcome some of the limitations of previous work, thirty native German speakers were recorded in a production study while interacting with a confederate (one of the experimenters) in a referential game, following experimental procedures typical of the field.
The participants had to verbally instruct the confederate to select a specified target object out of four objects presented on a screen.
The subject and confederate were seated at the opposite sides of a table, each facing one of two computer screens.
The participant and the experimenter could not see each other nor each others' screens.
Figure \ref{fig:procedural} shows the experimental procedure time-line.
After a familiarization phase, the subject first saw four colored objects in the top left, top right, bottom left, and bottom right corners of the screen.
One of the objects served as the target, another as the competitor, and the remaining two objects served as distractors.
Objects were referred to using noun phrases consisting of an adjective modifier denoting color and a modified object (e.g.~\emph{Gelbe Zitrone} `yellow lemon', \emph{Rote Gurke} `red cucumber', \emph{Rote Socken} `red socks').

@@

The two sentence prompts were used to create a focus contrast between the competitor and the target object.
Focused units denote the set of all (contextually relevant) alternatives \citep[e.g.][]{rooth1992theory}.
Concretely, a focus contrast marks one of more elements in a sentence as prominent, by different linguistic means depending on the language.
For instance, if the competitor and target objects differ but their color does not (e.g.~\emph{yellow banana} vs \emph{yellow tomato}), the noun is said to be in focus.
We call this the Noun Focus condition, or NF for short.
If the objects are the same but differ in color (e.g.~\emph{yellow banana} vs \emph{blue banana}), the color adjective is in focus (the Adjective Focus condition, AF).
If both the color and the object differ (e.g.~\emph{yellow banana} vs \emph{blue tomato}), then the whole noun phrase is in focus (the Adjective/Noun Focus condition, ANF).
The NF condition constituted the experimentally relevant condition, while the AF and ANF conditions acted as fillers.
Crucially, the color-object combinations in the Noun Focus (NF) condition were manipulated with respect to their typicality.
The combinations were either typical (e.g.~\emph{orange mandarin}), medium typical (e.g.~\emph{green tomato}), or atypical (e.g.~\emph{yellow cherry}), as established by a norming study that was conducted prior to the production experiment just described.\footnote{A detailed description of the norming and production studies from the \emph{Prosodic encoding of redundant referring expressions} project, which was given to the analysts with the data set, can be found in \texttt{methods\_norm\_prod.pdf} at (temporary link) \url{https://osf.io/5agn9/?view_only=3429acfd8eae4f0aafe2e17ad5b0983f}}
Each subject produced 15 critical trials (NF condition).
Each trial was repeated twice, yielding a total of 30 trials per participant and a grand-total of 900 (\(15 \times 2 \times 30\) participants) spoken utterances.

For the present study, N analysis teams will receive access to the entire data set generated by the production study.
The data set is constituted by audio recordings and annotation files in a format that is typical for the field.
The teams will be instructed to answer the following research question, using the provided data set: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}

@@

We estimate that this process, from the time of an in-principle acceptance of this Stage 1 Registered Report to the end of Phase 5, will take nine months.
The factor most likely to delay our time-line is the rate of completion of the original set of analyses by independent analysis teams.

The project OSF repository contains all the materials mentioned in this paper and can be accessed at LINK.\footnote{Temporary link: \url{https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f}.}
The main repository holds two OSF components, \texttt{Data} and \texttt{Questionnaires}, and it is linked to the GitHub repository.
The following sections report the criteria for sample size, data exclusions, data manipulations, and all the measures in the study.

@@

The analysis teams will register for participation and each of the analysts individually will answer a demographic and expertise questionnaire (\texttt{intake\_form.pdf}).\footnote{A PDF version of this and all other questionnaires are available in the repository's \texttt{Questionnaires} component, at \url{https://osf.io/h6z8w/?view_only=6faaae5feabe4eed800a7c142fcda06f}. The direct link to the intake form questionnaire is \url{https://osf.io/w7db5/?view_only=6faaae5feabe4eed800a7c142fcda06f}.}
The questionnaire collects information on the analysts current position and self-estimated breadth and level of statistical expertise and acoustic analysis skills.
We will then request that they answer the question: \emph{Do speakers acoustically modify utterances to signal atypical word combinations?}
To do so, they will be given the data generated by the experiment described in Section \protect\hyperlink{s:dataset}{The data set}.
The data will include the audio recordings with corresponding time-aligned transcriptions in the form of Praat TextGrid files.\footnote{These can be found in the \texttt{Data} component at \url{https://osf.io/3bmcp/?view_only=6faaae5feabe4eed800a7c142fcda06f}.}
